Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd

We use the statistical programming language R [cite R]... to perform data simulation, testing, cleaning, and analysis.

@@

### Outcome Variable
The outcome variable in this study is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S. Presidential Election based on polling data from key battleground states. This variable categorizes the election outcome into two distinct categories: a value of "Trump" indicates that Donald Trump is predicted to win, while a value of "Harris" denotes a predicted victory for Kamala Harris. The classification is derived from a conditioning statement that assigns the predicted winner based on the probability of Trump winning; if this probability exceeds 0.5, the predicted winner is labeled as Trump, otherwise, it is labeled as Harris. This binary outcome variable serves a crucial role in facilitating the analysis of voter support dynamics and the effectiveness of the predictive modeling employed in this paper.

By focusing on this dichotomy, we can glean insights into how polling data might translate into electoral outcomes, highlighting the competitive nature of the election in these pivotal states. The construction of the outcome variable not only enhances our understanding of voter preferences but also underscores the significance of binary classification in electoral forecasting, especially in states that could ultimately sway the overall election results.

### Predictive Variable
The predictive variable in this study is the polling support percentage for each candidate, specifically for Donald Trump and Kamala Harris, measured across various battleground states. This variable represents the predicted likelihood of each candidate winning in the specified states based on the polling data collected during the election cycle.
```{r}
#| label: tbl-outcomevar
#| fig-cap: "Average polling support percentages for Donald Trump and Kamala Harris across key battleground states. The table presents the average support levels, indicating Trump’s and Harris’s polling figures in states such as Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada. The comparative analysis highlights the varying levels of candidate support in these critical regions, providing insight into the electoral landscape ahead of the 2024 U.S. Presidential Election."
#| echo: false
#| warning: false
#| message: false

@@

```
@tbl-predictivevar highlights the average support levels for both Donald Trump and Kamala Harris across key battleground states, including Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada. The values reflect the polling percentages for each candidate, providing a straightforward comparison of their standings within these critical regions.

Understanding these polling percentages is essential as they offer insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election. The battleground states depicted in the table are particularly significant due to their historical tendency to swing between parties, making them crucial targets for campaign efforts. Analyzing the outcome variable through these polling percentages informs our predictive modeling and enhances our comprehension of the dynamics at play during the electoral process. This comparison not only contextualizes the current political climate but also underlines the importance of these battleground states in shaping the overall election results. As the campaign progresses, these insights will be invaluable for strategizing and understanding voter sentiment in these pivotal areas.

>>>>

The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race. For all the following variables mentioned, we collected observations for both of the candidates. We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris. We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’ and ‘state’. Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data quality are further explained in {additional data details}. The electorate support for each candidate is denoted as ‘pct’.

All of the variables mentioned above are used during the steps of data cleaning, simulation, testing, and analysis. We use the statistical programming language R [cite R]... to perform data simulation, testing, cleaning, and analysis.

@@

### Outcome Variable
The outcome variable in our model is ‘predicted_winner’, which is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S. Presidential Election based on polling data from key battleground states. This variable categorizes the election outcome into two distinct categories: a value of "Trump" indicates that Donald Trump is predicted to win, while a value of "Harris" denotes a predicted victory for Kamala Harris. The classification is derived from a conditioning statement that assigns the predicted winner based on the probability of Trump winning; if this probability exceeds 0.5, the predicted winner is labeled as Trump, otherwise, it is labeled as Harris.

The ‘predicted_winner’ variable in our model is a constructed binary indicator derived from the pct variable based on our model. The “pct” variable represents the percentage of support for each candidate, Harris and Trump, within each state. To create the “winner” variable, we compare the “pct” values: if Trump’s percentage of support in a given state is higher than Harris’s, the “winner” variable is assigned a value indicating Trump as the predicted winner for the given state; otherwise, it is assigned to Harris. This construction allows us to translate individual support levels from the raw data into a binary prediction of each state’s likely outcome.

### Predictive Variable
The predictive variables for this paper are each candidate’s support rate (‘pct’), which directly influences the “winner” variable. The ‘pct’ variable represents the percentage of the vote or support that each candidate received in the poll. Understanding these polling percentages is crucial, as they provide insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election. The battleground states presented in the table are particularly significant due to their historical tendency to swing between parties, making them vital focal points for campaign efforts. By examining the outcome variable through these polling percentages, we enhance our predictive modeling and gain a deeper understanding of the factors influencing the election. This comparison not only contextualizes the current political climate but also highlights the critical role these battleground states play in determining the overall election results. As the campaign unfolds, these insights will be invaluable for strategizing and gauging voter sentiment in these pivotal areas.

```{r}
#| label: tbl-outcomevar
#| fig-cap: "Average polling support percentages for Donald Trump and Kamala Harris across key battleground states. The table presents the average support levels, indicating Trump’s and Harris’s polling figures in states such as Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada. The comparative analysis highlights the varying levels of candidate support in these critical regions, providing insight into the electoral landscape ahead of the 2024 U.S. Presidential Election."
#| echo: false
#| warning: false
#| message: false

@@

```
@tbl-predictivevar highlights the average support levels for both Donald Trump and Kamala Harris across key battleground states, including Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada. The values reflect the polling percentages for each candidate, based on 'pct', providing a straightforward comparison of the predictive variables in each battleground state. It is important to note that since ‘pct’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%. In reality, as seen in Arizona, the combined ‘pct’ for Trump and Harris is 92.9%, with the remaining 7.1% likely reflecting support for other candidates in the race.
